---
title: "manuscript codes"
author: "Gloria"
format: html
editor: visual
execute: 
  warning: FALSE
---

## Libraries and Sources

```{r}
#| echo = FALSE
rm(list=ls())
#library(reshape2)
library(ggplot2)
library(bootstrap)
library(scales)
library("dplyr")
library(ggplot2)
library(vctrs)
library(tidyr)
library(tidyverse)
#library(shiny)
library(deSolve)
library(ggplot2)
#library(gridGraphics)   # Disable this after testing to avoid messing up Shiny
#library(shinyjs)
#library(shinythemes)
source("ODEModel.R")
source("ODEMosquitoParameters.R")
source("ODEAuxiliaryFunctions.R")
source("ODEControlMeasuresParameters.R")
source("ODETransmissionParameters.R")
source("ODEInterventions.R")
#source("ODEInterventions - FirstPass.R")
source("ODEModelOutput.R")
source("FeedingCycle.R")
source("multiplot.R")
```

Get parameter for a specific specie

Different start time for LAR and BIO will crush at the max time entered - to be fixed Turn on required specie

```{r,echo = FALSE}
#| echo: FALSE

MOSQUITO_PARAMETERS_1 = getAnGambiaeParameters()
# MOSQUITO_PARAMETERS_2 = getAnArabiensisParameters()
# MOSQUITO_PARAMETERS_3 = getAnFunestusParameters()

# simulation runs per day - enter the end time
INITIAL_MODELRUNTIME_VALUE   = 365

#Enter the coverage for a specific intervetion
# To turn it on  - enter the required time which is < the max model run time
# To turn it off - enter INITIAL_MODELRUNTIME_VALUE + 1

###************************************************************

#Source reduction, coverage value, time it is on
INITIAL_SRE_COVERAGE = 0.0
INITIAL_SRE_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#LArvaciding, coverage value, time it is on
INITIAL_LAR_COVERAGE = 0.0
INITIAL_LAR_TIME     = INITIAL_MODELRUNTIME_VALUE + 1 #75 

#Biological, coverage value, time it is on
INITIAL_BIO_COVERAGE = 0.0
INITIAL_BIO_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#ATSB, coverage value, time it is on
INITIAL_ATSB_COVERAGE = 0.0
INITIAL_ATSB_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Space Spraying, coverage value, time it is on
INITIAL_SSP_COVERAGE = .00
INITIAL_SSP_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Odor Traps, coverage value, time it is on
INITIAL_OBT_COVERAGE = .00
INITIAL_OBT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#LLINs, coverage value, time it is on
INITIAL_ITN_COVERAGE = 0.3
INITIAL_ITN_TIME     = 0
#IRS
INITIAL_IRS_COVERAGE = 0.0
INITIAL_IRS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#House modification
INITIAL_HOU_COVERAGE = 0.0
INITIAL_HOU_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Personal Protection measure
INITIAL_PPM_COVERAGE = 0.0
INITIAL_PPM_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Systemic
INITIAL_ECS_COVERAGE = 0.0
INITIAL_ECS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Topical
INITIAL_ECT_COVERAGE = 0.0
INITIAL_ECT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Resting and Ovipositing - OviTraps -assuming same coverage for ATSB and SSP
INITIAL_OVI_COVERAGE = 0.0
INITIAL_OVI_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

## Get intervetions parameters - LLINs for testing
#INTERVENTION_PARAMETERS = getInterventionsParameters(ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME)

## Get intervetions parameters
INTERVENTION_PARAMETERS = getInterventionsParameters(
  #Source Reduction
  SREcov=INITIAL_SRE_COVERAGE,time_SRE_on=INITIAL_SRE_TIME,
  #Larvaciding
  LARcov=INITIAL_LAR_COVERAGE,time_LAR_on=INITIAL_LAR_TIME,
  #Biological Control
  BIOcov=INITIAL_BIO_COVERAGE,time_BIO_on=INITIAL_BIO_TIME,
  
  #ATSB
  ATSBcov=INITIAL_ATSB_COVERAGE,time_ATSB_on=INITIAL_ATSB_TIME,
  #Space Spraying
  SSPcov=INITIAL_SSP_COVERAGE,time_SSP_on=INITIAL_SSP_TIME,
  #Odor Traps
  OBTcov=INITIAL_OBT_COVERAGE,time_OBT_on=INITIAL_OBT_TIME,
  #LLINs
  ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME,
  #IRS
  IRScov=INITIAL_IRS_COVERAGE,time_IRS_on=INITIAL_IRS_TIME,
  #House Modification
  HOUcov=INITIAL_HOU_COVERAGE,time_HOU_on=INITIAL_HOU_TIME,
  #Personal Protection Measure
  PPMcov=INITIAL_PPM_COVERAGE,time_PPM_on=INITIAL_PPM_TIME,
  #Cattle - Systemic
  ECScov=INITIAL_ECS_COVERAGE,time_ECS_on=INITIAL_ECS_TIME,
  #Cattle - topical
  ECTcov=INITIAL_ECT_COVERAGE,time_ECT_on=INITIAL_ECT_TIME,
  #Resting & Ovipositing - Ovitraps --same for ATSB, SSP
  OVIcov=INITIAL_OVI_COVERAGE,time_OVI_on=INITIAL_OVI_TIME)

#Pass in interventions parameter with updated coverage and the specie

theta <<- getTheta(interventionParameters=INTERVENTION_PARAMETERS, 
                   speciesSpecificParameters=MOSQUITO_PARAMETERS_1)


## Initialize the model
initState <<- calculateInitialState(theta)

IVM_traj <<- runODE(INITIAL_MODELRUNTIME_VALUE,1,initState,theta,"daspk")

IVM_traj <- IVM_traj[-c(1,2), ]
```

Run the model - eventually produce different IVM_traj for different scenarios, eg., 1 intervention, 2 inter, etc

```{r,echo=FALSE}
#| echo: FALSE


MOSQUITO_PARAMETERS_1 = getAnGambiaeParameters()
# MOSQUITO_PARAMETERS_2 = getAnArabiensisParameters()
# MOSQUITO_PARAMETERS_3 = getAnFunestusParameters()

# simulation runs per day - enter the end time
INITIAL_MODELRUNTIME_VALUE   = 365

#Enter the coverage for a specific intervetion
# To turn it on  - enter the required time which is < the max model run time
# To turn it off - enter INITIAL_MODELRUNTIME_VALUE + 1

###************************************************************

#Source reduction, coverage value, time it is on
INITIAL_SRE_COVERAGE = 0.0
INITIAL_SRE_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#LArvaciding, coverage value, time it is on
INITIAL_LAR_COVERAGE = 0.6
INITIAL_LAR_TIME     = 150 

#Biological, coverage value, time it is on
INITIAL_BIO_COVERAGE = 0.0
INITIAL_BIO_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#ATSB, coverage value, time it is on
INITIAL_ATSB_COVERAGE = 0.0
INITIAL_ATSB_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Space Spraying, coverage value, time it is on
INITIAL_SSP_COVERAGE = .00
INITIAL_SSP_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Odor Traps, coverage value, time it is on
INITIAL_OBT_COVERAGE = .00
INITIAL_OBT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#LLINs, coverage value, time it is on
INITIAL_ITN_COVERAGE = 0.3
INITIAL_ITN_TIME     = 0
#IRS
INITIAL_IRS_COVERAGE = 0.0
INITIAL_IRS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#House modification
INITIAL_HOU_COVERAGE = 0.0
INITIAL_HOU_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Personal Protection measure
INITIAL_PPM_COVERAGE = 0.0
INITIAL_PPM_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Systemic
INITIAL_ECS_COVERAGE = 0.0
INITIAL_ECS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Topical
INITIAL_ECT_COVERAGE = 0.0
INITIAL_ECT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Resting and Ovipositing - OviTraps -assuming same coverage for ATSB and SSP
INITIAL_OVI_COVERAGE = 0.0
INITIAL_OVI_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

## Get intervetions parameters - LLINs for testing
#INTERVENTION_PARAMETERS = getInterventionsParameters(ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME)

## Get intervetions parameters
INTERVENTION_PARAMETERS = getInterventionsParameters(
  #Source Reduction
  SREcov=INITIAL_SRE_COVERAGE,time_SRE_on=INITIAL_SRE_TIME,
  #Larvaciding
  LARcov=INITIAL_LAR_COVERAGE,time_LAR_on=INITIAL_LAR_TIME,
  #Biological Control
  BIOcov=INITIAL_BIO_COVERAGE,time_BIO_on=INITIAL_BIO_TIME,
  
  #ATSB
  ATSBcov=INITIAL_ATSB_COVERAGE,time_ATSB_on=INITIAL_ATSB_TIME,
  #Space Spraying
  SSPcov=INITIAL_SSP_COVERAGE,time_SSP_on=INITIAL_SSP_TIME,
  #Odor Traps
  OBTcov=INITIAL_OBT_COVERAGE,time_OBT_on=INITIAL_OBT_TIME,
  #LLINs
  ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME,
  #IRS
  IRScov=INITIAL_IRS_COVERAGE,time_IRS_on=INITIAL_IRS_TIME,
  #House Modification
  HOUcov=INITIAL_HOU_COVERAGE,time_HOU_on=INITIAL_HOU_TIME,
  #Personal Protection Measure
  PPMcov=INITIAL_PPM_COVERAGE,time_PPM_on=INITIAL_PPM_TIME,
  #Cattle - Systemic
  ECScov=INITIAL_ECS_COVERAGE,time_ECS_on=INITIAL_ECS_TIME,
  #Cattle - topical
  ECTcov=INITIAL_ECT_COVERAGE,time_ECT_on=INITIAL_ECT_TIME,
  #Resting & Ovipositing - Ovitraps --same for ATSB, SSP
  OVIcov=INITIAL_OVI_COVERAGE,time_OVI_on=INITIAL_OVI_TIME)

#Pass in interventions parameter with updated coverage and the specie

theta <<- getTheta(interventionParameters=INTERVENTION_PARAMETERS, 
                   speciesSpecificParameters=MOSQUITO_PARAMETERS_1)


## Initialize the model
initState <<- calculateInitialState(theta)

IVM_traj_2 <<- runODE(INITIAL_MODELRUNTIME_VALUE,1,initState,theta,"daspk")

IVM_traj_2<- IVM_traj_2[-c(1,2), ]
```

```{r,echo = FALSE}
#| echo: FALSE


MOSQUITO_PARAMETERS_1 = getAnGambiaeParameters()

# simulation runs per day - enter the end time
INITIAL_MODELRUNTIME_VALUE   = 365

#Enter the coverage for a specific intervetion
# To turn it on  - enter the required time which is < the max model run time
# To turn it off - enter INITIAL_MODELRUNTIME_VALUE + 1

###************************************************************

#Source reduction, coverage value, time it is on
INITIAL_SRE_COVERAGE = 0.0
INITIAL_SRE_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#LArvaciding, coverage value, time it is on
INITIAL_LAR_COVERAGE = 0.0
INITIAL_LAR_TIME     = 0 

#Biological, coverage value, time it is on
INITIAL_BIO_COVERAGE = 0.0
INITIAL_BIO_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#ATSB, coverage value, time it is on
INITIAL_ATSB_COVERAGE = 0.0
INITIAL_ATSB_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Space Spraying, coverage value, time it is on
INITIAL_SSP_COVERAGE = .00
INITIAL_SSP_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Odor Traps, coverage value, time it is on
INITIAL_OBT_COVERAGE = .00
INITIAL_OBT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#LLINs, coverage value, time it is on
INITIAL_ITN_COVERAGE = 0.1
INITIAL_ITN_TIME     = 0
#IRS
INITIAL_IRS_COVERAGE = 0.0
INITIAL_IRS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#House modification
INITIAL_HOU_COVERAGE = 0.0
INITIAL_HOU_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Personal Protection measure
INITIAL_PPM_COVERAGE = 0.0
INITIAL_PPM_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Systemic
INITIAL_ECS_COVERAGE = 0.0
INITIAL_ECS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Topical
INITIAL_ECT_COVERAGE = 0.0
INITIAL_ECT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Resting and Ovipositing - OviTraps -assuming same coverage for ATSB and SSP
INITIAL_OVI_COVERAGE = 0.0
INITIAL_OVI_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

## Get intervetions parameters - LLINs for testing
#INTERVENTION_PARAMETERS = getInterventionsParameters(ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME)

## Get intervetions parameters
INTERVENTION_PARAMETERS = getInterventionsParameters(
  #Source Reduction
  SREcov=INITIAL_SRE_COVERAGE,time_SRE_on=INITIAL_SRE_TIME,
  #Larvaciding
  LARcov=INITIAL_LAR_COVERAGE,time_LAR_on=INITIAL_LAR_TIME,
  #Biological Control
  BIOcov=INITIAL_BIO_COVERAGE,time_BIO_on=INITIAL_BIO_TIME,
  
  #ATSB
  ATSBcov=INITIAL_ATSB_COVERAGE,time_ATSB_on=INITIAL_ATSB_TIME,
  #Space Spraying
  SSPcov=INITIAL_SSP_COVERAGE,time_SSP_on=INITIAL_SSP_TIME,
  #Odor Traps
  OBTcov=INITIAL_OBT_COVERAGE,time_OBT_on=INITIAL_OBT_TIME,
  #LLINs
  ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME,
  #IRS
  IRScov=INITIAL_IRS_COVERAGE,time_IRS_on=INITIAL_IRS_TIME,
  #House Modification
  HOUcov=INITIAL_HOU_COVERAGE,time_HOU_on=INITIAL_HOU_TIME,
  #Personal Protection Measure
  PPMcov=INITIAL_PPM_COVERAGE,time_PPM_on=INITIAL_PPM_TIME,
  #Cattle - Systemic
  ECScov=INITIAL_ECS_COVERAGE,time_ECS_on=INITIAL_ECS_TIME,
  #Cattle - topical
  ECTcov=INITIAL_ECT_COVERAGE,time_ECT_on=INITIAL_ECT_TIME,
  #Resting & Ovipositing - Ovitraps --same for ATSB, SSP
  OVIcov=INITIAL_OVI_COVERAGE,time_OVI_on=INITIAL_OVI_TIME)

#Pass in interventions parameter with updated coverage and the specie

theta <<- getTheta(interventionParameters=INTERVENTION_PARAMETERS, 
                   speciesSpecificParameters=MOSQUITO_PARAMETERS_1)


## Initialize the model
initState <<- calculateInitialState(theta)

IVM_traj_3 <<- runODE(INITIAL_MODELRUNTIME_VALUE,1,initState,theta,"daspk")

IVM_traj_3<- IVM_traj_3[-c(1,2), ]
```

```{r,echo = FALSE}
#| echo: FALSE


MOSQUITO_PARAMETERS_1 = getAnGambiaeParameters()
# MOSQUITO_PARAMETERS_2 = getAnArabiensisParameters()
# MOSQUITO_PARAMETERS_3 = getAnFunestusParameters()

# simulation runs per day - enter the end time
INITIAL_MODELRUNTIME_VALUE   = 365

#Enter the coverage for a specific intervetion
# To turn it on  - enter the required time which is < the max model run time
# To turn it off - enter INITIAL_MODELRUNTIME_VALUE + 1

###************************************************************

#Source reduction, coverage value, time it is on
INITIAL_SRE_COVERAGE = 0.0
INITIAL_SRE_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#LArvaciding, coverage value, time it is on
INITIAL_LAR_COVERAGE = 0.8
INITIAL_LAR_TIME     = 150 

#Biological, coverage value, time it is on
INITIAL_BIO_COVERAGE = 0.0
INITIAL_BIO_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#ATSB, coverage value, time it is on
INITIAL_ATSB_COVERAGE = 0.0
INITIAL_ATSB_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Space Spraying, coverage value, time it is on
INITIAL_SSP_COVERAGE = .00
INITIAL_SSP_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Odor Traps, coverage value, time it is on
INITIAL_OBT_COVERAGE = .00
INITIAL_OBT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#LLINs, coverage value, time it is on
INITIAL_ITN_COVERAGE = 0.1
INITIAL_ITN_TIME     = 0
#IRS
INITIAL_IRS_COVERAGE = 0.0
INITIAL_IRS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#House modification
INITIAL_HOU_COVERAGE = 0.0
INITIAL_HOU_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Personal Protection measure
INITIAL_PPM_COVERAGE = 0.0
INITIAL_PPM_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Systemic
INITIAL_ECS_COVERAGE = 0.0
INITIAL_ECS_TIME     = INITIAL_MODELRUNTIME_VALUE + 1
#Cattle - Topical
INITIAL_ECT_COVERAGE = 0.0
INITIAL_ECT_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

#Resting and Ovipositing - OviTraps -assuming same coverage for ATSB and SSP
INITIAL_OVI_COVERAGE = 0.0
INITIAL_OVI_TIME     = INITIAL_MODELRUNTIME_VALUE + 1

## Get intervetions parameters - LLINs for testing
#INTERVENTION_PARAMETERS = getInterventionsParameters(ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME)

## Get intervetions parameters
INTERVENTION_PARAMETERS = getInterventionsParameters(
  #Source Reduction
  SREcov=INITIAL_SRE_COVERAGE,time_SRE_on=INITIAL_SRE_TIME,
  #Larvaciding
  LARcov=INITIAL_LAR_COVERAGE,time_LAR_on=INITIAL_LAR_TIME,
  #Biological Control
  BIOcov=INITIAL_BIO_COVERAGE,time_BIO_on=INITIAL_BIO_TIME,
  
  #ATSB
  ATSBcov=INITIAL_ATSB_COVERAGE,time_ATSB_on=INITIAL_ATSB_TIME,
  #Space Spraying
  SSPcov=INITIAL_SSP_COVERAGE,time_SSP_on=INITIAL_SSP_TIME,
  #Odor Traps
  OBTcov=INITIAL_OBT_COVERAGE,time_OBT_on=INITIAL_OBT_TIME,
  #LLINs
  ITNcov=INITIAL_ITN_COVERAGE,time_ITN_on=INITIAL_ITN_TIME,
  #IRS
  IRScov=INITIAL_IRS_COVERAGE,time_IRS_on=INITIAL_IRS_TIME,
  #House Modification
  HOUcov=INITIAL_HOU_COVERAGE,time_HOU_on=INITIAL_HOU_TIME,
  #Personal Protection Measure
  PPMcov=INITIAL_PPM_COVERAGE,time_PPM_on=INITIAL_PPM_TIME,
  #Cattle - Systemic
  ECScov=INITIAL_ECS_COVERAGE,time_ECS_on=INITIAL_ECS_TIME,
  #Cattle - topical
  ECTcov=INITIAL_ECT_COVERAGE,time_ECT_on=INITIAL_ECT_TIME,
  #Resting & Ovipositing - Ovitraps --same for ATSB, SSP
  OVIcov=INITIAL_OVI_COVERAGE,time_OVI_on=INITIAL_OVI_TIME)

#Pass in interventions parameter with updated coverage and the specie

theta <<- getTheta(interventionParameters=INTERVENTION_PARAMETERS, 
                   speciesSpecificParameters=MOSQUITO_PARAMETERS_1)


## Initialize the model
initState <<- calculateInitialState(theta)

IVM_traj_4 <<- runODE(INITIAL_MODELRUNTIME_VALUE,1,initState,theta,"daspk")

IVM_traj_4<- IVM_traj_4[-c(1,2), ]
```

# RESULTS SECTION

```{r,echo = FALSE}
#| echo: FALSE
#plotTrajectory(IVM_traj)
dig1<-plotTrajectory(IVM_traj_2)
dig2<-plotTrajectoryimmature(IVM_traj_2)

# ggsave("mature_plot.png", plot = dig1,width = 10, height = 6, units = "in", dpi = 300, device = "png")
# 
# ggsave("immature_plot.png", plot = dig2,width = 10, height = 6, units = "in", dpi = 300, device = "png")



DIGNEW<-ggplot(IVM_traj_2)+
  geom_line(aes(x=time,y=LL),color="green",size=1.75)+
  theme_classic()+
  labs(x="Time(days)",y="Late Instar")+
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 90, size = 12),  # Adjust the size for better visibility
    axis.text.y = element_text(size = 12),  # Adjust the size for better visibility
    axis.title.x = element_text(size = 20),  # Adjust the size for better visibility
    axis.title.y = element_text(size = 20),
    legend.position = "right",
    legend.title = element_text( size = 20),
    legend.text = element_text(size = 20))

# ggsave("larval_plot.png", plot = DIGNEW,width = 10, height = 6, units = "in", dpi = 300, device = "png")


```

```{r,echo = FALSE}
df_stages <- IVM_traj_2 %>%
  select(EL, LL, PL, SV, EV, IV)

# Convert to long format
df_long <- df_stages %>%
  pivot_longer(cols = everything(),
               names_to = "Stage",
               values_to = "Number")
# Fix the order of stages
df_long$Stage <- factor(df_long$Stage,
                        levels = c("EL", "LL", "PL", "SV", "EV", "IV"))




di4<-ggplot(df_long, aes(x = Stage, y = log(Number), fill = Stage)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(
    limits = c(0, NA),   # force y-axis to start at 0
    labels = scales::comma
  ) +
  labs(x = "Mosquito Stage", y = "Number of Mosquitoes") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 90, size = 12),  # Adjust the size for better visibility
    axis.text.y = element_text(size = 12),  # Adjust the size for better visibility
    axis.title.x = element_text(size = 20),  # Adjust the size for better visibility
    axis.title.y = element_text(size = 20),
    legend.position = "right",
    legend.title = element_text( size = 20),
    legend.text = element_text(size = 20)) +
  scale_fill_discrete(labels = c("EL" = "early instar", "LL" = "late instar", "PL" = "pupae", "SV" = "susceptible mosquito", "EV" = "exposed mosquito", "IV" = "infected mosquito"))
# ggsave("bar_plot.png", plot = di4,width = 10, height = 6, units = "in", dpi = 300, device = "png")
```

# fig 5 combined

```{r,echo = FALSE}
library(gridExtra)
r1<-grid.arrange(dig1,dig2,DIGNEW,di4)
# ggsave("combined_plot.png", plot = r1,width = 10, height = 6, units = "in", dpi = 300, device = "png")

r1
```

# infected mosquito reduction

```{r,echo = FALSE}
library(boot)
# Calculate the total number of infected cases before and after intervention
total_cases_before <- sum(IVM_traj_3$IV)
total_cases_after <- sum(IVM_traj_4$IV)

# Calculate the percentage reduction
percentage_reduction <- ((total_cases_before - total_cases_after) / total_cases_before) * 100

percentage_reduction

```

# eir reduction

```{r,echo = FALSE}
library(boot)
# Calculate the total number of infected cases before and after intervention
total_cases_before <- sum(IVM_traj$EIR)
total_cases_after <- sum(IVM_traj_2$EIR)

# Calculate the percentage reduction
percentage_reduction <- ((total_cases_before - total_cases_after) / total_cases_before) * 100

percentage_reduction
```

# fig6.b - mdensity plot

```{r,echo = FALSE}
#| echo: FALSE

md_plot <- ggplot() +
  theme_grey(base_size = 18, base_family = "") +
  ylim(0, NA) +
  geom_line(data=IVM_traj_3, aes(x=time, y=Mdensity,colour="Before Larv"),size=1.75) + 
  geom_line(data=IVM_traj_4, aes(x=time, y=Mdensity,colour="After Larv"),size=1.75) +
  labs(x = "days", y = "Mosquito Density", color = "Scenario")+
  theme_classic()+
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 90, size = 12),  # Adjust the size for better visibility
    axis.text.y = element_text(size = 12),  # Adjust the size for better visibility
    axis.title.x = element_text(size = 20),  # Adjust the size for better visibility
    axis.title.y = element_text(size = 20),
    legend.position = "right",
    legend.title = element_text( size = 20),
    legend.text = element_text(size = 20)
  )
md_plot
# ggsave("md_plot.png", plot = md_plot, width = 10, height = 6, units = "in", dpi = 300)
```

# reduction percent in mdensity

```{r,echo = FALSE}
library(boot)
# Calculate the total number of infected cases before and after intervention
total_cases_before <- sum(IVM_traj$Mdensity)
total_cases_after <- sum(IVM_traj_2$Mdensity)

# Calculate the percentage reduction
percentage_reduction <- ((total_cases_before - total_cases_after) / total_cases_before) * 100

percentage_reduction
```

# figure 1 monthly data collected

```{r,echo = FALSE}
monthly_sum <- read.csv("monthly_data_collection.csv")


# Extract year and create a color column
dfn <- monthly_sum %>%
  mutate(Color = ifelse(month < "2020-07-01", "Control", "Intervention"))

monthly_sum1 <- monthly_sum %>%
  mutate(
    month = as.Date(month),  # if it is already ISO-like "YYYY-MM-DD"
    Color = ifelse(month < as.Date("2020-07-01"), "Control", "Intervention"),
    month_label = format(month, "%b-%y"),
    month = as.character(month)
  )


# Keep only every 2nd month for x-axis ticks to reduce crowding
tick_breaks <- monthly_sum1$month[seq(1, nrow(monthly_sum1), by = 2)]
tick_labels <- monthly_sum1$month_label[seq(1, nrow(monthly_sum1), by = 2)]

# Create the plot
fig1ms <- ggplot(monthly_sum1, aes(x = month, y = total, group = 1)) +
  geom_line(aes(color = Color), size = 1.75) +
  geom_point() +
  
  # Dashed vertical line separating the two phases
  geom_vline(xintercept = "2020-07-01", linetype = "dashed") +
  
  # Label on the dashed line to clarify what it marks
  annotate("text",
           x = "2020-07-01",
           y = max(monthly_sum1$total) * 0.95,
           label = "Larviciding\nintroduced",
           hjust = -0.1,
           size = 4,
           fontface = "plain") +
  
  
  # Simplified x-axis tick labels (every 2nd month, Mon-YY format)
  scale_x_discrete(breaks = tick_breaks, labels = tick_labels) +
  
  # Regular (non-italic) legend labels, corrected color mapping
  scale_color_manual(values = c("Control" = "red", "Intervention" = "skyblue"),
                     labels = c("Control", "Intervention")) +
  
  labs(title = "",
       x = "Time (Months)",
       y = "Number of Mosquitoes Collected",
       color = " ") +
  
  theme_classic() +
  theme(
    # X-axis: larger font, still angled for readability
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    # Legend: plain (not italic) font
    legend.text = element_text(face = "plain"),
    legend.title = element_blank()
  )
# ggsave("figure1ms_plot.png", plot = fig1ms, width = 10, height = 6, units = "in", dpi = 300)


 

```

# figure 2 - rufiji map

```{r,echo = FALSE}
# Install packages if missing
#install.packages(c("geodata", "sf", "ggplot2", "dplyr", "cowplot"))
library(ggplot2)
library(cowplot)
library(sf)
library(dplyr)
library(geodata)

# Load GADM data
tanzania <- gadm(country = "TZA", level = 1, path = tempdir())
tanzania_sf <- st_as_sf(tanzania)
pwani <- tanzania_sf %>% filter(NAME_1 == "Pwani")

tanzania_l2 <- gadm(country = "TZA", level = 2, path = tempdir())
tanzania_l2_sf <- st_as_sf(tanzania_l2)
pwani_districts <- tanzania_l2_sf %>% filter(NAME_1 == "Pwani")
rufiji <- pwani_districts %>% filter(NAME_2 == "Rufiji")

# MAP 1: Tanzania
map_tz <- ggplot() +
  geom_sf(data = tanzania_sf, fill = "grey80", color = "white") +
  geom_sf(data = pwani, fill = "orange", color = "black", size = 0.4) +
  theme_void()

# MAP 2: Pwani
map_pwani <- ggplot() +
  geom_sf(data = pwani_districts, fill = "grey80", color = "white", size = 0.3) +
  geom_sf(data = rufiji, fill = "orange", color = "black", size = 0.6) +
  theme_void()

# MAP 3: Rufiji
map_rufiji <- ggplot() +
  geom_sf(data = rufiji, fill = "orange", color = "black", size = 0.4) +
  theme_void()

# Combine maps
combined <- ggdraw() +
  draw_plot(map_tz, x = 0, y = 0, width = 0.4, height = 1) +
  draw_plot(map_pwani, x = 0.45, y = 0.4, width = 0.3, height = 0.5) +
  draw_plot(map_rufiji, x = 0.8, y = 0.55, width = 0.15, height = 0.3)

# Now add curved arrows using ggplotGrob and annotation_custom
library(grid)

# Tanzania -> Pwani
arrow1 <- linesGrob(
  x = c(0.38, 0.45),
  y = c(0.65, 0.65),
  arrow = arrow(length = unit(0.15, "inches")),
  gp = gpar(col = "black", lwd = 2)
)

# Pwani -> Rufiji
arrow2 <- linesGrob(
  x = c(0.72, 0.8),
  y = c(0.65, 0.7),
  arrow = arrow(length = unit(0.15, "inches")),
  gp = gpar(col = "black", lwd = 2)
)

# Draw everything together
grid.newpage()
grid.draw(combined)
grid.draw(arrow1)
grid.draw(arrow2)

```
